Progress Report

### 2015-12-22 Report Dec 2015/Report.tex
Report Dec initial

Performance analysis of GPIC algorithm is provided in \cite{radji2009interference}, here we briefly conclude the main results in \cite{radji2009interference}, at the receiving side, the frame error is defined as there is at least one erroneous symbol in the estimation of transmitted symbol vector. Let $P_{e}$ denotes the average frame error probability. The diversity order, denoted by $d$ is defined as the slope of the $P_{e}$ in log-scale in high SNR region, which is given by\cite{zheng2003diversity}
\begin{equation}
d=-\lim_{SNR\to\infty}\frac{\log{P_{e}}}{\log(SNR)},
\label{diversity}
\end{equation}
Let $P_{e}^{ML}$ denotes the average frame error probability of MLD, and $P_{e2}=\mathbb{E}_{\mathbf{H}_{2}, \mathbf{s}_{2}}(Pr(\hat{\mathbf{s}}^{k_{1}}_{2}\neq \mathbf{s}_{2}|\mathbf{H}_{2}, \mathbf{s}_{2}))$, where $P_{r}(A)$ denotes the probability that a event A occurs, $\hat{\mathbf{s}}_{1}^{k_{1}}=\mathbf{s}_{1}$, therefore $P_{e2}$ is the average frame error probability of the LD/SICs for the sub system in which the interference from the users that transmitted $\mathbf{s}_{1}$ are perfectly cancelled from the observation $\mathbf{y}$. given by
\begin{equation}
\mathbf{y}^{k_{1}}=\mathbf{y}-\mathbf{H}_{1}\hat{\mathbf{s}}_{1}^{k_{1}},
\label{perfect IC}
\end{equation} 
because $\hat{\mathbf{s}}^{k1}_{1}=\mathbf{s}_{1}$, we have
\begin{equation}
\mathbf{y}^{k1}=\mathbf{H}_{2}\mathbf{s}_{2}+\mathbf{n},
\label{sub MIMO system}
\end{equation} 
The average frame error probability of GPIC algorithm $P_{et}$ is bounded by 
\begin{equation}
\max(P^{ML}_{e}, P_{e2})\leq P_{et}\leq P^{ML}_{e}+P_{e2}, 
\label{average frame error probability of GPIC}
\end{equation}
in (\ref{average frame error probability of GPIC}), the diversity order of MLD is $N_{r}$, $P_{e2}$ is the average frame error probability of sub optimal LD/SICs, let $d_{2}$ denotes the diversity order of LD/SICs, based on (\ref{diversity}), we have 
\begin{equation}
\lim_{SNR\to\infty}P^{ML}_{e}\propto SNR^{-N_{r}},
\label{MLD diversity}
\end{equation}
\begin{equation}
\lim_{SNR\to\infty}P_{e2}\propto SNR^{-d_{2}},
\label{LD diversity}
\end{equation}
Let $d_{t}$ denotes the diversity order of GPIC algorithm, now consider the following conditions
\begin{itemize}
\item $d_{2}>N_{r}$,
\begin{equation}
\lim_{SNR\to\infty}\frac{P_{e2}}{P^{ML}_{e}}=SNR^{N_{r}-d_{2}}=0,
\label{condition1sub1}
\end{equation}
based on (\ref{average frame error probability of GPIC}), we have
\begin{equation}
\lim_{SNR\to\infty}\frac{P_{et}}{P^{ML}_{e}}=1,
\label{condition1sub2}
\end{equation}
thus $d_{t}=N_{r}$.
\item $d_2=N_{r}$,
we have 
\begin{equation}
\lim_{SNR\to\infty}P_{et}\propto SNR^{-N_{r}},
\label{condition2sub1}
\end{equation}
thus $d_{t}=N_{r}$.
\item $d_{2}<N_{r}$,
\begin{equation}
\lim_{SNR\to\infty}\frac{P^{ML}_{e}}{P_{e2}}=SNR^{d_{2}-N_{r}}=0,
\label{condition3sub1}
\end{equation}
based on (\ref{average frame error probability of GPIC}), we have  
\begin{equation}
\lim_{SNR\to\infty}P_{et}=P_{e2}\propto SNR^{-d_{2}},
\label{condition3sub2}
\end{equation}
thus 
$d_{t}=d_{2}$.
\end{itemize} 
Therefore when $d_{2}\geq N_{r}$, GPIC algorithm can achieve ML performance asymptotically, when $d_{2}<N_{r}$, the diversity order that GPIC can achieve at high SNR region is equal to that of LD/SICs, in \cite{radji2009interference}, the authors employ diversity maximization selection (DMS) scheme for channel partition, which is proposed in \cite{zhang2006diversity}. The optimum diversity order can be guaranteed with low complexity.

To elaborate further, for a given number of antennas chosen at the channel partition stage $N$, there is $N_{u}=\left[\begin{array}{l}N_{t}\\N\end{array}\right]$ possible combinations of $[\mathbf{H}_{1}, \mathbf{H}_{2}]$, according to DMS principle, the maximum diversity order of LD/SIC is achieved by choosing the subset $\mathbf{H}_{2}$ that has the strongest weakest substream, in a sense of post processing SNR. Therefore, when using MMSE/SIC as the sub optimal detector, the subset $\mathbf{H}^{opt}_{2}$ selected based on DMS principle should be\cite{zhang2006diversity} 
\begin{IEEEeqnarray}[\relax]{l}
\mathbf{H}^{opt}_{2}=\mathbf{H}^{p}_{2},\\
p=\min_{j=1,2,\ldots, N_{u}}\theta_{j},\\
\theta_{j}=\max_{k=1,2,\ldots, N_{t}-N}((\mathbf{H}_{2}^{j})^{H}\mathbf{H}_{2}^{j}+SNR^{-1}\mathbf{I})_{kk}^{-1},\quad \mathbf{H}_{2}^{j}\in[\mathbf{H}_{2}^{1}, \mathbf{H}_{2}^{2}, \ldots, \mathbf{H}_{2}^{N_{u}}];
\label{DMS principle}
\end{IEEEeqnarray}   
Where $\mathbf{A}_{kk}$ denotes the $k$th diagonal component of matrix $\mathbf{A}$. The diversity order of LD/SIC detector after DMS channel partition process is given by 
$d_{2DMS}=(N+1)(N_{r}-N_{t}+N+1)$. In \cite{radji2009interference}, the authors derive the minimum number of antennas $N_{min}$ at channel partition stage that can guarantee $d_{2DMS}\geq N_{r}$, which is given by 
\begin{equation}
N_{min}=\lceil \sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+N_{r}}-\frac{N_{r}-N_{t}}{2}-1\rceil,
\label{Nmin DMS}
\end{equation}
where $\lceil\alpha\rceil$ denotes the minimum integer no less than $\alpha$.
\subsection{Minimum Achievable Diversity: How many antennas do we need for LS-MIMO?}
In LS-MIMO V-BLAS systems, the maximum diversity order is $N_{r}$, which is extremely large under the condition that there is tens to hundreds of receive antennas. On the one hand, diversity order is achieved at high SNR region, however, in practical, such a high diversity order is not necessary, because the SNR region of interest is the range in which the bit error rate (BER) or symbol error rate (SER) can be $10^{-5}$ to $10^{-7}$. On the other hand, the computational complexity to guarantee ML performance by DMS channel partition is excessive in LS-MIMO, for example, when $N_{r}=N_{t}$, based on (\ref{Nmin DMS}), the $N_{min}=\lceil\sqrt{N_{r}}-1\rceil$, the number of the symbol vector candidates in the list generated by exhaustive search for $\mathbf{s}_{1}$ is $M^{N_{min}}$.

Therefore, here we consider minimum achievable diversity (MAD) principle, which sacrifices redundant diversity gain in LS-MIMO, in order to reduce the complexity. Let $d_{2MAD}$ denotes the diversity order of LD/SIC detectors, $\tilde{N}_{min}$ denotes the minimum number of antennas chosen by MAD principle. Our goal is to guarantee the overall diversity is no less than a given minimum diversity while keep $\tilde{N}_{min}$ small. Let $1\leq g<N_{r}$ denote the given minimum achievable diversity order, based on (\ref{condition3sub2}), we have
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
 if \quad d_{2MAD}<N_{r}\quad and\quad d_{2MAD}\geq g\\
  d_{t}=d_{2MAD}\geq g.
\end{IEEEeqnarray}
Therefore we can guarantee that the overall diversity order of GPIC algorithm is no less than $g$, in order to derive the minimum number of antennas required based on MAD principle, we define function
\begin{IEEEeqnarray}[\relax]{l}
f(N)=d_{2MAD}-g=(N+1)(N_{r}-N_{t}+N+1)-g,\label{diversity function}
\end{IEEEeqnarray}    
where $N\in [1,2,\ldots, N_{t}]$, our goal is to find minimum $N$, denoted by $\tilde{N}_{min}$, that satisfy $f(N)\geq 0$. The two zeros points of quadratic function $f(N)$ are 
\begin{IEEEeqnarray}[\relax]{l}
N_{1}=-\sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1,\\
N_{2}=\sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1
\label{zeros points}
\end{IEEEeqnarray}
when $N<N_{1}$ or $N>N_{2}$, $f(N)\geq 0$, obviously $N_{1}<0$, there is no effective $N$ exits that satisfy the former condition, thus, we have 
\begin{equation}
\tilde{N}_{min}=\lceil \sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1\rceil
\label{Nmin MAD}
\end{equation}
\section{Theoretical Analysis of Channel Hardening}
\subsection{Preliminary}
Orthogonality deficiency $\phi_{od}$ measures  how orthogonal a matrix is \cite{ma2008performance}, which is defined by
\begin{equation}
\phi_{od}=1-\frac{det(\mathbf{W})}{\prod_{i=1}^{N_{t}}||\mathbf{h}_{i}||^{2}},
\label{OD}
\end{equation}
where $\mathbf{W}=\mathbf{H}^{H}\mathbf{H}$ denotes Wishart matrix, $\mathbf{h}_{i}$ denotes the $i$ th column of $\mathbf{H}$, $det(\cdot)$ denotes determinant operation, $||\cdot||$ denotes 2-norm operation. Based on Hadamard's inequality $\prod_{i=1}^{N_{t}}||\mathbf{h}_{i}||\geq det(\mathbf{H})$, we have $0\leq\phi_{od}\leq 1$, if $\mathbf{H}$ is singular, $\phi_{od}=1$, if $\mathbf{H}$ is orthogonal, then $\phi_{od}=0$.

@@

For sake of simplicity, we define orthogonality measure $\phi_{om}$, which is given by 
\begin{equation}
\phi_{om}=\frac{det(\mathbf{W})}{\prod_{i=1}^{N_{t}}||\mathbf{h}_{i}||^{2}},
\label{OM}
\end{equation}
$0\leq \phi_{om}\leq 1$, if $\phi_{om}$ is closer to 1, $\mathbf{H}$ is more closer to an orthogonal matrix. 
\subsection{Introduction to Channel Hardening Phenomenon}
\subsection{Logarithmic Expectation of Orthogonality Measure}
Based on (\ref{OM}), The logarithm of $\phi_{om}$ can be written as 
\begin{equation}
\ln(\phi_{om})=\ln(det(\mathbf{W}))-\sum_{i=1}^{N_{t}}\ln(||\mathbf{h}_{i}||^{2}),
\label{log OM}
\end{equation}
Taking expectation of (\ref{log OM}), we have 
\begin{equation}
\mathbb{E}[\ln(\phi_{om})]=\mathbb{E}[\ln(det(\mathbf{W}))]-\sum_{i=1}^{N_{t}}\mathbb{E}[\ln(||\mathbf{h}_{i}||^{2})].
\label{log expectation OM}
\end{equation}
First we consider the first summand on the right hand of (\ref{log expectation OM}) $\mathbb{E}[\ln(det(\mathbf{W}))]$. Let $\mathbb{C}W_{m}(n, \Sigma)$ denotes complex Wishart distribution, which is the joint distribution of sample covariance matrix from multivariate complex Gaussian random variable\cite{goodman1963statistical}. $n$ is the degree of freedom and $\Sigma\in \mathbb{C}^{m\times m}$ is the covariance matrix. Let $\Pi_{i}$ denotes the $i$th row of $\mathbf{H}$. $\Pi_{i}$ is a $N_{t}$-variate complex Gaussian random variable so does $\Pi_{i}^{H}$. Therefore $\mathbf{W}=\mathbf{H}^{H}\mathbf{H}=\sum_{i=1}^{N_{r}}\Pi_{i}^{H}\Pi_{i}\sim \mathbb{C}W_{N_{t}}(N_{r}, \mathbf{I})$. The logarithmic expectation of $det(\mathbf{W})$ is
\begin{equation}
\mathbb{E}[\ln(det(\mathbf{W}))]=\sum_{i=1}^{N_{t}}\psi(N_{r}-i+1),
\label{log expectation of wishart}
\end{equation}
where $\psi(n)$ denotes Digamma function, which is given by\cite{papoulis2002probability}
\begin{equation}
\psi(n)=\frac{\Gamma^{'}(n)}{\Gamma(n)},
\label{Digmma function}
\end{equation}
$\Gamma(n)$ denotes Gamma function.
Proof: see Appendix \ref{log expectation of OM}.

@@

Based on (\ref{log expectation of wishart}) and (\ref{sum of log expectation gamma}), the logarithmic expectation of orthogonality measure $\phi_{om}$ is 
\begin{equation}
\mathbb{E}[\ln(\phi_{om})]=\sum_{i=1}^{N_{t}}[\psi(N_{r}-i+1)-\psi(N_{r})],
\label{final log expectation OM}
\end{equation}
\subsection{Probability Density Function of Orthogonality Measure}
\section{Simplified Channel Partition based on Orthogonality Measure}
\section{Computer Simulations}
\subsection{Full Loaded Systems}
\subsection{Systems with Different Loading Factors}

\section{Conclusion}
The conclusion goes here.

>>>>

Performance analysis of GPIC algorithm is provided in \cite{radji2009interference}, here we briefly conclude the main results in \cite{radji2009interference}, at the receiving side, the frame error is defined as there is at least one erroneous symbol in the estimation of transmitted symbol vector. Let $P_{e}$ denotes the average frame error probability. The diversity order, denoted by $d$ is defined as the slope of the $P_{e}$ in log-scale in high SNR region, which is given by\cite{zheng2003diversity}
\begin{equation}
d=-\lim_{SNR\to\infty}\frac{\log{P_{e}}}{\log(SNR)},
\label{diversity}
\end{equation}
Let $P_{e}^{ML}$ denotes the average frame error probability of MLD, and $P_{e2}=\mathbb{E}_{\mathbf{H}_{2}, \mathbf{s}_{2}}(Pr(\hat{\mathbf{s}}^{k_{1}}_{2}\neq \mathbf{s}_{2}|\mathbf{H}_{2}, \mathbf{s}_{2}))$, where $P_{r}(A)$ denotes the probability that a event A occurs, $\hat{\mathbf{s}}_{1}^{k_{1}}=\mathbf{s}_{1}$, therefore $P_{e2}$ is the average frame error probability of the LD/SICs for the sub system in which the interference from the users that transmitted $\mathbf{s}_{1}$ are perfectly cancelled from the observation $\mathbf{y}$. given by
\begin{equation}
\mathbf{y}^{k_{1}}=\mathbf{y}-\mathbf{H}_{1}\hat{\mathbf{s}}_{1}^{k_{1}},
\label{perfect IC}
\end{equation} 
because $\hat{\mathbf{s}}^{k1}_{1}=\mathbf{s}_{1}$, we have
\begin{equation}
\mathbf{y}^{k1}=\mathbf{H}_{2}\mathbf{s}_{2}+\mathbf{n},
\label{sub MIMO system}
\end{equation} 
The average frame error probability of GPIC algorithm $P_{et}$ is bounded by 
\begin{equation}
\max(P^{ML}_{e}, P_{e2})\leq P_{et}\leq P^{ML}_{e}+P_{e2}, 
\label{average frame error probability of GPIC}
\end{equation}
in (\ref{average frame error probability of GPIC}), the diversity order of MLD is $N_{r}$, $P_{e2}$ is the average frame error probability of sub optimal LD/SICs, let $d_{2}$ denotes the diversity order of LD/SICs, based on (\ref{diversity}), we have 
\begin{equation}
\lim_{SNR\to\infty}P^{ML}_{e}\propto SNR^{-N_{r}},
\label{MLD diversity}
\end{equation}
\begin{equation}
\lim_{SNR\to\infty}P_{e2}\propto SNR^{-d_{2}},
\label{LD diversity}
\end{equation}
Let $d_{t}$ denotes the diversity order of GPIC algorithm, now consider the following conditions
\begin{itemize}
\item $d_{2}>N_{r}$,
\begin{equation}
\lim_{SNR\to\infty}\frac{P_{e2}}{P^{ML}_{e}}=SNR^{N_{r}-d_{2}}=0,
\label{condition1sub1}
\end{equation}
based on (\ref{average frame error probability of GPIC}), we have
\begin{equation}
\lim_{SNR\to\infty}\frac{P_{et}}{P^{ML}_{e}}=1,
\label{condition1sub2}
\end{equation}
thus $d_{t}=N_{r}$.
\item $d_2=N_{r}$,
we have 
\begin{equation}
\lim_{SNR\to\infty}P_{et}\propto SNR^{-N_{r}},
\label{condition2sub1}
\end{equation}
thus $d_{t}=N_{r}$.
\item $d_{2}<N_{r}$,
\begin{equation}
\lim_{SNR\to\infty}\frac{P^{ML}_{e}}{P_{e2}}=SNR^{d_{2}-N_{r}}=0,
\label{condition3sub1}
\end{equation}
based on (\ref{average frame error probability of GPIC}), we have  
\begin{equation}
\lim_{SNR\to\infty}P_{et}=P_{e2}\propto SNR^{-d_{2}},
\label{condition3sub2}
\end{equation}
thus 
$d_{t}=d_{2}$.
\end{itemize} 
Therefore when $d_{2}\geq N_{r}$, GPIC algorithm can achieve ML performance asymptotically, when $d_{2}<N_{r}$, the diversity order that GPIC can achieve at high SNR region is equal to that of LD/SICs, in \cite{radji2009interference}, the authors employ diversity maximization selection (MDS) scheme for channel partition, which is proposed in \cite{zhang2006diversity}. The optimal diversity order can be guaranteed with low complexity for conventional MIMO systems.

To elaborate further, for a given number of antennas chosen at the channel partition stage $N$, there is $N_{u}={N_{t}\choose N}$ possible combinations of $[\mathbf{H}_{1}, \mathbf{H}_{2}]$, according to MDS principle, the maximum diversity order of LD/SIC is achieved by choosing the subset $\mathbf{H}_{2}$ that has the strongest weakest substream, in a sense of post processing SNR. Therefore, when using MMSE/SIC as the sub optimal detector, the subset $\mathbf{H}^{opt}_{2}$ selected based on DMS principle should be\cite{zhang2006diversity} 
\begin{IEEEeqnarray}[\relax]{l}
\mathbf{H}^{opt}_{2}=\mathbf{H}^{p}_{2},\\
p=\min_{j=1,2,\ldots, N_{u}}\theta_{j},\\
\theta_{j}=\max_{k=1,2,\ldots, N_{t}-N}((\mathbf{H}_{2}^{j})^{H}\mathbf{H}_{2}^{j}+SNR^{-1}\mathbf{I})_{kk}^{-1},\quad \mathbf{H}_{2}^{j}\in[\mathbf{H}_{2}^{1}, \mathbf{H}_{2}^{2}, \ldots, \mathbf{H}_{2}^{N_{u}}];
\label{DMS principle}
\end{IEEEeqnarray}   
Where $\mathbf{A}_{kk}$ denotes the $k$th diagonal component of matrix $\mathbf{A}$. The diversity order of LD/SIC detector after DMS channel partition process is given by 
$d_{2DMS}=(N+1)(N_{r}-N_{t}+N+1)$. In \cite{radji2009interference}, the authors derive the minimum number of antennas $N_{min}$ at channel partition stage that can guarantee $d_{2DMS}\geq N_{r}$, which is given by 
\begin{equation}
N_{min}=\lceil \sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+N_{r}}-\frac{N_{r}-N_{t}}{2}-1\rceil,
\label{Nmin DMS}
\end{equation}
where $\lceil\alpha\rceil$ denotes the minimum integer no less than $\alpha$.
\subsection{Minimum Achievable Diversity: How many antennas do we need for LS-MIMO?}
In LS-MIMO V-BLAS systems, the maximum diversity order is $N_{r}$, which is extremely large under the condition that there is tens to hundreds of receive antennas. On the one hand, diversity order is achieved at high SNR region, however, in practical, such a high diversity order is not necessary, because the SNR region of interest is the range in which the bit error rate (BER) or symbol error rate (SER) can be $10^{-5}$ to $10^{-7}$. On the other hand, the computational complexity to guarantee ML performance by DMS channel partition is excessive in LS-MIMO, for example, when $N_{r}=N_{t}$, based on (\ref{Nmin DMS}), the $N_{min}=\lceil\sqrt{N_{r}}-1\rceil$, the number of the symbol vector candidates in the list generated by exhaustive search for $\mathbf{s}_{1}$ is $M^{N_{min}}$.

Therefore, here we consider minimum achievable diversity (MAD) principle, which sacrifices redundant diversity gain in LS-MIMO, in order to reduce the complexity. Let $d_{2MAD}$ denotes the diversity order of LD/SIC detectors, $\tilde{N}_{min}$ denotes the minimum number of antennas chosen by MAD principle. Our goal is to guarantee the overall diversity is no less than a given minimum diversity while keep $\tilde{N}_{min}$ small. Let $1\leq g<N_{r}$ denote the given minimum achievable diversity order, based on (\ref{condition3sub2}), we have
\begin{IEEEeqnarray}[\relax]{l}
\nonumber
 if \quad d_{2MAD}<N_{r}\quad and\quad d_{2MAD}\geq g\\
  d_{t}=d_{2MAD}\geq g.
\end{IEEEeqnarray}
Therefore we can guarantee that the overall diversity order of GPIC algorithm is no less than $g$, in order to derive the minimum number of antennas required based on MAD principle, we define function
\begin{IEEEeqnarray}[\relax]{l}
f(N)=d_{2MAD}-g=(N+1)(N_{r}-N_{t}+N+1)-g,\label{diversity function}
\end{IEEEeqnarray}    
where $N\in [1,2,\ldots, N_{t}]$, our goal is to find minimum $N$, denoted by $\tilde{N}_{min}$, that satisfy $f(N)\geq 0$. The two zeros points of quadratic function $f(N)$ are 
\begin{IEEEeqnarray}[\relax]{l}
N_{1}=-\sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1,\\
N_{2}=\sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1
\label{zeros points}
\end{IEEEeqnarray}
when $N\leq N_{1}$ or $N\geq N_{2}$, $f(N)\geq 0$, obviously $N_{1}<0$, there is no feasible $N$ exits that can make the former condition satisfied. Thus, we have 
\begin{equation}
\tilde{N}_{min}=\lceil \sqrt{\frac{(N_{r}-N_{t})^{2}}{4}+g}-\frac{N_{r}-N_{t}}{2}-1\rceil
\label{Nmin MAD}
\end{equation}
\subsection{Computer Simulations}
In this section, computer simulation results for the bit error rate (BER) performances of MAD principle based GPIC algorithm are presented. As comparisons, Sel-MMSE-OSIC algorithm in \cite{radji2009interference}, which is based on maximum diversity selection (MDS) principle is also considered. The software testbed is build by C, compiled by GCC compiler version 4.9.2 on 64 bit Debian (release 8.2) Linux systems, the experiments are performed on two desktops, one with Intel I5-4th generation CPU, quad cores, 3.2GHz clock rate and the other one with Intel I7-5th generation CPU, six cores, 3.5GHz clock rate.

We consider uncoded complex spatial multiplexing LS-MIMO systems, For each receive SNR point, the BER are calculated with minimum $10^{5}$ independent channel realizations achieved and minimum 300 symbol errors accumulated. The modulation schemes considered are rectangular M-QAM(4-QAM and 16-QAM) with Gray code labeling. In each channel realization, $N_{t}$ mutually independent randomly generated bit sequences are modulated to complex symbols. $N_{t}$ complex symbols are transmitted over randomly generated Rayleigh flat fading channel, each channel matrix component is CSCG random variable with unit variance. The receive symbol vector are corrupted by AWGN.

For sake of simplicity, we use MAD-n to present the results for minimum achievable diversity selection, where $n$ denotes the minimum achievable diversity. The minimum number of antennas selected by channel partition for a given minimum achievable diversity is given by (\ref{Nmin MAD}). Maximum diversity selection (MDS) is used to present the Sel-MMSE-OSIC that can achieve ML diversity asymptotically. The minimum number of antennas selected by channel partition is given by (\ref{Nmin DMS}).

First, we consider $16\times 16$ system with $4$-QAM modulation, Table.\ref{t16MIMO4QAM} shows the schemes considered, the corresponding results are presented in Fig.\ref{f16MIMO4QAM}, 
\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{$16\times 16$ system $4$-QAM}
\label{t16MIMO4QAM}
\centering
\begin{tabular}{|c|c|c|}
\hline
Scheme&number of antennas selected&Asymptotic diversity\\
\hline 
MDS&3&16\\
\hline
MAD-9&2&9\\
\hline
MAD-4&1&4\\
\hline
\end{tabular}
\end{table}

As shown in Fig.\ref{f16MIMO4QAM}, compared with MDS which can achieve ML performance asymptotically, MAD-9 has a small performance loss, while MAD-4's performance is inferior, at BER $10^{-5}$, the MAD-9 is about $0.8$dB worse than MDS, MAD-4 is about $1.8$dB worse than MDS, let $\kappa(\cdot)$ denote the number of list candidate generated by GPIC algorithm\cite{radji2009interference}, from Table.\ref{t16MIMO4QAM}, 
we have $\frac{\kappa(MAD-9)}{\kappa(DMS)}=1/4$ and $\frac{\kappa(MAD-4)}{\kappa(DMS)}=1/16$.

\begin{figure}[htb]
\centering
%\includegraphics[width=\textwidth]{16MIMO4QAM.eps}
\def\svgwidth{\columnwidth}
\input{16MIMO4QAM.pdf_tex}
\caption{$16\times 16$ system $4$-QAM}
\label{f16MIMO4QAM}
\end{figure}

The results for $16\times 16$ system with $16$-QAM modulation is shown in Fig.\ref{f16MIMO16QAM}, the information of the schemes considered is as the same as that in Talbe.\ref{t16MIMO4QAM}, From Fig.\ref{f16MIMO16QAM}, we can see that similar to the case of $4$-QAM system, MAD-9 has a very small performance loss comparing to DMS, at BER $2\times 10^{-5}$, MAD-9 is about $1.1$dB worse than MDS, while MAD-4 has about $3$dB  performance loss comparing to DMS.

Furthermore,comparing to $4$-QAM system, the reduction of the number of list candidates provided by MAD scheme is more significant in $16$-QAM, $\frac{\kappa(MAD-9)}{\kappa(DMS)}=1/16$ and $\frac{\kappa(MAD-4)}{\kappa(DMS)}=1/256$.
\begin{figure}[htb]
\centering
\def\svgwidth{\columnwidth}
\input{16MIMO16QAM.pdf_tex}
\caption{$16\times 16$ system $16$-QAM}
\label{f16MIMO16QAM}
\end{figure}

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{$32\times 32$ system $4$-QAM}
\label{t32MIMO4QAM}
\centering
\begin{tabular}{|c|c|c|}
\hline
Scheme&number of antennas selected&Asymptotic diversity\\
\hline 
MAD-4&1&4\\
\hline
MAD-9&2&9\\
\hline
MAD-16&3&16\\
\hline
\end{tabular}
\end{table}

\begin{figure}[htb]
\centering
\def\svgwidth{\columnwidth}
\input{32MIMO4QAM.pdf_tex}
\caption{$32\times 32$ system $4$-QAM}
\label{f32MIMO4QAM}
\end{figure}
Now we consider the performance of MAD in large systems, Fig.\ref{f32MIMO4QAM} shows the results for $32\times 32$ system with $4$-QAM. Table.\ref{t32MIMO4QAM} shows the information of the schemes, due to the time limitation, we just provide the results for MAD. As we can see in Fig\ref{f32MIMO4QAM}, although the asymptotic diversity of MAD-16 is much higher than that of MAD-9, at the BER region of interest, their performance difference is negligible, at BER=$3\times 10^{-4}$, the performance of MAD-9 is about $0.3$dB worse comparing to MAD-16, at BER=$3\times 10^{-5}$, the performance loss of MAD-9 is about $0.5$dB comparing to MAD-16.

\section{Theoretical Analysis of Channel Hardening Phenomenon}
\subsection{Preliminary}
Orthogonality deficiency $\phi_{od}$ measures  how orthogonal a matrix is \cite{ma2008performance}, which is defined by
\begin{equation}
\phi_{od}=1-\frac{det(\mathbf{W})}{\prod_{i=1}^{N_{t}}||\mathbf{h}_{i}||^{2}},
\label{OD}
\end{equation}
where $\mathbf{W}=\mathbf{H}^{H}\mathbf{H}$ denotes Wishart matrix, $\mathbf{h}_{i}$ denotes the $i$ th column of $\mathbf{H}$, $det(\cdot)$ denotes determinant operation, $||\cdot||$ denotes 2-norm operation. Based on Hadamard's inequality $\prod_{i=1}^{N_{t}}||\mathbf{h}_{i}||\geq det(\mathbf{H})$, we have $0\leq\phi_{od}\leq 1$, if $\mathbf{H}$ is singular, $\phi_{od}=1$, if $\mathbf{H}$ is orthogonal, then $\phi_{od}=0$.

@@

For sake of simplicity, we define orthogonality measure $\phi_{om}$, which is given by 
\begin{equation}
\phi_{om}=\frac{det(\mathbf{W})}{\prod_{i=1}^{N_{t}}||\mathbf{h}_{i}||^{2}},
\label{OM}
\end{equation}
$0\leq \phi_{om}\leq 1$, $\forall \mathbf{H}$, if $\phi_{om}$ is closer to 1, $\mathbf{H}$ is more closer to an orthogonal matrix. 
%\subsection{Introduction to Channel Hardening Phenomenon}
\subsection{Logarithmic Expectation of Orthogonality Measure}
Based on (\ref{OM}), The logarithm of $\phi_{om}$ can be written as 
\begin{equation}
\ln(\phi_{om})=\ln(det(\mathbf{W}))-\sum_{i=1}^{N_{t}}\ln(||\mathbf{h}_{i}||^{2}),
\label{log OM}
\end{equation}
Taking expectation of (\ref{log OM}), we have 
\begin{equation}
\mathbb{E}[\ln(\phi_{om})]=\mathbb{E}[\ln(det(\mathbf{W}))]-\sum_{i=1}^{N_{t}}\mathbb{E}[\ln(||\mathbf{h}_{i}||^{2})].
\label{log expectation OM}
\end{equation}
First we consider the first summand on the right hand of (\ref{log expectation OM}) $\mathbb{E}[\ln(det(\mathbf{W}))]$. Let $\mathbb{C}W_{m}(n, \Sigma)$ denotes complex Wishart distribution, which is the joint distribution of sample covariance matrix from multivariate complex Gaussian random variable\cite{goodman1963statistical}. $n$ is the degree of freedom and $\Sigma\in \mathbb{C}^{m\times m}$ is the covariance matrix. Let $\Pi_{i}$ denotes the $i$th row of $\mathbf{H}$. $\Pi_{i}$ is a $N_{t}$-variate complex Gaussian random variable, so does $\Pi_{i}^{H}$. Therefore $\mathbf{W}=\mathbf{H}^{H}\mathbf{H}=\sum_{i=1}^{N_{r}}\Pi_{i}^{H}\Pi_{i}\sim \mathbb{C}W_{N_{t}}(N_{r}, \mathbf{I})$. The logarithmic expectation of $det(\mathbf{W})$ is
\begin{equation}
\mathbb{E}[\ln(det(\mathbf{W}))]=\sum_{i=1}^{N_{t}}\psi(N_{r}-i+1),
\label{log expectation of wishart}
\end{equation}
where $\psi(n)$ denotes Digamma function, which is given by\cite{papoulis2002probability}
\begin{equation}
\psi(n)=\frac{\Gamma^{'}(n)}{\Gamma(n)},
\label{Digmma function}
\end{equation}
$\Gamma(n)$ denotes Gamma function.
Proof: see Appendix \ref{log expectation of OM}.

@@

Based on (\ref{log expectation of wishart}) and (\ref{sum of log expectation gamma}), the logarithmic expectation of orthogonality measure $\phi_{om}$ is 
\begin{equation}
\mathbb{E}[\ln(\phi_{om})]=\sum_{i=1}^{N_{t}}[\psi(N_{r}-i+1)-\psi(N_{r})],
\label{final log expectation OM}
\end{equation}
\subsection{Probability Density Function of Orthogonality Measure}
In this section, we derive the probability distribution of $\phi_{om}$. First, let's consider an alternative form of (\ref{OM}), do QR factorization to $\mathbf{H}$, $\mathbf{H}=\mathbf{Q}\mathbf{R}$\cite{watkins2004fundamentals}, where $\mathbf{Q}\in\mathbb{C}^{N_{r}\times N_{t}}$ is a unitary matrix and $\mathbf{R}\in\mathbb{C}^{N_{t}\times N_{t}}$ is an upper triangular matrix. Thus $\mathbf{W}=\mathbf{H}^{H}\mathbf{H}=\mathbf{R}^{H}\mathbf{R}$. Let $r_{ji}, j\leq i$ denote the component of $\mathbf{R}$ at $j$th row and $i$th column, $r^{*}_{ji}$ denotes the conjugate of $r_{ji}$, we have
\begin{equation}
det(\mathbf{W})=det(\mathbf{R}^{H}\mathbf{R})=det(\mathbf{R}^{H})det(\mathbf{R})\\=\prod_{i=1}^{N_{t}}r_{ii}^{*}\prod_{i=1}^{N_{t}}r_{ii}=\prod_{i=1}^{N_{t}}|r_{ii}|^{2},
\label{alternative wishart}
\end{equation}
Furthermore, let $\mathbf{R}_{i}$ denote the $i$th column of $\mathbf{R}$, $\mathbf{W}_{ii}$ denote the $i$th diagonal component of $\mathbf{W}$, we have
\begin{equation}
\mathbf{W}_{ii}=||\mathbf{h}_{i}||^{2}=||\mathbf{R}_{i}||^{2}=\sum_{j<i}|r_{ji}|^{2}+|r_{ii}|^{2},
\label{alternative wishart diagonal}
\end{equation}
based on (\ref{alternative wishart}) and (\ref{alternative wishart diagonal}), (\ref{OM}) can be rewritten as 
\begin{equation}
\phi_{om}=\prod_{i=1}^{N_{t}}(\frac{|r_{ii}|^{2}}{|r_{ii}|^{2}+\sum_{j<i}|r_{ji}|^{2}}).
\label{alternative OM}
\end{equation}
